Self Regulation Ontology Retest Data Report

### 2017-11-06 code/SRO_Report.Rmd
prep for meeting

```{r warning=FALSE, message=FALSE}
lit_review = lit_review %>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)]),
         raw1_fit0 = grepl('raw', variable_type),
         type = as.character(type)) %>%
  mutate(task_group = gsub("_", " ", task_group),
         var = gsub("_", " ", var)) %>%
  arrange(task_group, raw1_fit0, var) %>%
  mutate(task_group = ifelse(task_group == "psychological refractory period two choices", "psychological refractory period", ifelse(task_group == "angling risk task always sunny", "angling risk task",task_group))) %>%
  mutate(task_group = gsub("survey", "", task_group)) %>%
    select(-measure_description, -reference)
```

@@

```{r echo=FALSE}
#Get variables of interest
tmp1 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables.csv')
tmp2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_noDDM.csv')
tmp3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_EZ.csv')
retest_report_vars = c(names(tmp1), names(tmp2), names(tmp3))
retest_report_vars = unique(retest_report_vars)
lit_rev_vars = unique(lit_review$dv)[which(unique(lit_review$dv) %in% retest_report_vars == FALSE)]
retest_report_vars = c(retest_report_vars, lit_rev_vars)
rm(tmp1, tmp2, tmp3, lit_rev_vars)
```

@@

## Reliability metrics 
For correlations we report Spearman's $\rho$'s only.  
Based on [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) ICC(3,k) does not take in to account within subject differences between two time points (i.e. the fixed effect of time/systematic error). Thus, it is well approximated by Pearson's r and subject to similar criticisms. [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) suggests reporting at least this systematic error effect size if one chooses to report with ICC(3,k). Based on his conclusions here I report:  
- ICC(3,k): As Dave clarified this ranges from 1 to -1/(number of repeated measures -1) so in our case this range would be [-1, 1]; larger values would mean that the two scores of a subject for a given measure are more similar to each other than they are to scores of other people  
- partial $\eta^2$ for time ($SS_{time}/SS_{within}$): effect size of time   
- SEM ($\sqrt(MS_{error})$): standard error of measurement; the smaller the better

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(label_1 %in% c('difference', 'EZ', 'hddm') ==FALSE & label_2 %in% c('difference', 'EZ', 'hddm') == FALSE) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  group_by(task_name) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            num_measures = n(),
            num_trials = unique(num_all_trials)) %>%
  arrange(-median_icc, -median_spearman)

@@

```{r warning=FALSE, message=FALSE}
summary(lm(median_icc ~ num_trials, data = tmp))
```

Does number of measures that went in to this reliability estimate have a significant effect on the average ICC of (mostly) raw measures for all trials from a task? No. (same is true for Spearman)

```{r warning=FALSE, message=FALSE}
summary(lm(median_icc ~ num_measures, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(ddm_task == 1) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.', remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  drop_na()
```

@@

```{r warning=FALSE, message=FALSE}
tmp %>%
  group_by(label_1,label_2, label_3) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            n_vars = n()) %>%
  arrange(-median_icc, -median_spearman) %>%
  datatable() %>%
  formatRound(columns=c('median_spearman', 'median_icc'), digits=3)
```

@@

```{r warning=FALSE, message=FALSE}
measure_labels %>%
  select(dv, task, variable_type, ddm_task, num_all_trials)%>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)])) %>%
  filter(task == "task") %>%
  select(-var, -task) %>%
  left_join(rel_df[,c("dv", "icc")], by="dv")%>%
  # filter(ddm_task == 1) %>%
  separate(variable_type, c("overall_difference", "raw_EZ_hddm", "param"), extra="merge") %>%
  mutate(raw_EZ_hddm = ifelse(ddm_task == 0, "other", raw_EZ_hddm),
         param = ifelse(ddm_task == 0, "other", param)) %>%
  na.omit() %>%
  filter(overall_difference!='condition') %>%
  ggplot(aes(x=factor(raw_EZ_hddm, levels = c("raw", "EZ", "hddm", "other"), labels=c("Raw", "EZ-diffusion", "Hierarchical diffusion", "Other")),y=icc,fill=param))+
  geom_boxplot(position = position_dodge(width=0.75))+
  facet_wrap(~factor(overall_difference, levels=c("overall", "difference"), labels=c("Overall", "Difference")))+
  theme_bw()+
  xlab("")+
  ylab("ICC")+
  theme(legend.title = element_blank())

@@

```{r}
summary(lm(icc ~ label_1,tmp))
```

When all trials are used the drift rates and rt's have higher reliabilities compared to accuracies.

```{r}
summary(lm(icc~label_3,tmp[tmp$label_1=="overall",]))
```

@@

ggplotly(tmp %>%
  ggplot(aes(days_btw, mean_diff))+
  geom_point()+
  theme_bw()+
  xlab("Days between completion")+
  ylab("Mean standardized difference between two time points")+
  geom_smooth(method="lm"))
```

@@

```{r}
boot_df <- read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/input/bootstrap_merged.csv')

@@

```{r warning=FALSE, message=FALSE}
tmp = rel_df %>%
  mutate(var_subs_pct = var_subs/(var_subs+var_ind+var_resid)*100,
         var_ind_pct = var_ind/(var_subs+var_ind+var_resid)*100, 
         var_resid_pct = var_resid/(var_subs+var_ind+var_resid)*100) %>%
  select(dv, task, var_subs_pct, var_ind_pct, var_resid_pct) %>%
  # gather(key, value, -dv, -task) %>%
  mutate(#key = factor(key, levels = c("var_subs_pct", "var_ind_pct", "var_resid_pct")),
         dv = factor(dv, levels = dv[order(task)])) %>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)])) %>%
  arrange(task_group, var_subs_pct) %>%
  mutate(rank = row_number()) %>%
  arrange(task, task_group, rank) %>%
  gather(key, value, -dv, -task_group, -var, -task, -rank) %>%
  ungroup()%>%
  # separate(var, c("var"), sep="\\.",remove=TRUE,extra="drop") %>%
  mutate(task_group = gsub("_", " ", task_group),
         var = gsub("_", " ", var),
         var=gsub(".ReflogTr", "", var),
         var=gsub(".logTr", "", var)) %>%
  mutate(task_group = ifelse(task_group == "psychological refractory period two choices", "psychological refractory period", ifelse(task_group == "angling risk task always sunny", "angling risk task",task_group))) %>%
  mutate(task_group = gsub("survey", "", task_group)) %>%
  filter(task=="task",
         !grepl("EZ|hddm", dv))%>%
  arrange(task_group, rank)
labels = tmp %>%
  distinct(dv, .keep_all=T)

@@

ggsave('Variance_Breakdown_Plot.jpg', plot = p3, device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures", width = 24, height = 20, units = "in")
```

@@

```{r}
tmp %>%
  group_by(task, key) %>%
  ggplot(aes(task, value, fill=key))+
  geom_boxplot()+
  theme_bw()+
  theme(legend.title = element_blank())+
  scale_fill_discrete(breaks = c("var_subs_pct", "var_ind_pct", "var_resid_pct"),
                    labels = c("Variance between individuals",
                               "Variance between sessions",
                               "Error variance"),
                    h=c(100,500))+
  xlab("")+
  ylab("Percent")

ggsave('Variance_Breakdown_Plot_Summary.jpg', device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures", width = 6, height = 5, units = "in")
```

>>>>

```{r warning=FALSE, message=FALSE}
lit_review = lit_review %>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)]),
         type = as.character(type)) %>%
  mutate(task_group = gsub("_", " ", task_group),
         var = gsub("_", " ", var)) %>%
  arrange(task_group, raw_fit, var) %>%
  mutate(task_group = ifelse(task_group == "psychological refractory period two choices", "psychological refractory period", ifelse(task_group == "angling risk task always sunny", "angling risk task",task_group))) %>%
  mutate(task_group = gsub("survey", "", task_group)) %>%
    select(-measure_description, -reference)
```

@@

```{r echo=FALSE}
#Get variables of interest from Ian's release
tmp1 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables.csv')
tmp2 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_noDDM.csv')
tmp3 <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_10-08-2017/meaningful_variables_EZ.csv')
retest_report_vars = c(names(tmp1), names(tmp2), names(tmp3))
retest_report_vars = unique(retest_report_vars)
lit_rev_vars = unique(lit_review$dv)[which(unique(lit_review$dv) %in% retest_report_vars == FALSE)]
retest_report_vars = c(retest_report_vars, lit_rev_vars)
rm(tmp1, tmp2, tmp3, lit_rev_vars)
```

@@

## Reliability metrics

Based on [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) ICC(3,k) does not take in to account within subject differences between two time points (i.e. the fixed effect of time/systematic error). Thus, it is well approximated by Pearson's r and subject to similar criticisms. [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) suggests reporting at least this systematic error effect size if one chooses to report with ICC(3,k). Based on his conclusions here I report:  
- ICC(3,k): As Dave clarified this ranges from 1 to -1/(number of repeated measures -1) so in our case this range would be [-1, 1]; larger values would mean that the two scores of a subject for a given measure are more similar to each other than they are to scores of other people  
- partial $\eta^2$ for time ($SS_{time}/SS_{within}$): effect size of time   
- SEM ($\sqrt(MS_{error})$): standard error of measurement; the smaller the better

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  filter(overall_difference != 'difference' & raw_fit %in% c('EZ', 'hddm') == FALSE)%>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  group_by(task_name) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            num_measures = n(),
            num_trials = unique(num_all_trials)) %>%
  arrange(-median_icc, -median_spearman)

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  filter(overall_difference != 'difference' & raw_fit %in% c('EZ', 'hddm') == FALSE)%>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2)

summary(lm(icc ~ num_all_trials, data = tmp))
```

```{r warning=FALSE, message=FALSE}
measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  filter(overall_difference != 'difference' & raw_fit %in% c('EZ', 'hddm') == FALSE)%>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  ggplot(aes(num_all_trials, icc))+
  geom_point()+
  geom_smooth(method="lm")+
  theme_bw()+
  xlab("Number of trials")+
  ylab("ICC")
```

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  filter(ddm_task == 1) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.', remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  drop_na()
```

@@

```{r warning=FALSE, message=FALSE}
tmp %>%
  group_by(overall_difference, raw_fit, rt_acc) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            n_vars = n()) %>%
  arrange(-median_icc, -median_spearman) %>%
  datatable() %>%
  formatRound(columns=c('median_spearman', 'median_icc'), digits=3)
```

@@

```{r warning=FALSE, message=FALSE}
measure_labels %>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)])) %>%
  filter(task == "task") %>%
  select(-var, -task) %>%
  left_join(rel_df[,c("dv", "icc")], by="dv")%>%
  na.omit() %>%
  filter(overall_difference!='condition',
         raw_fit != 'fit',
         rt_acc != 'other') %>%
  ggplot(aes(x=factor(raw_fit, levels = c("raw", "EZ", "hddm"), labels=c("Raw", "EZ-diffusion", "Hierarchical diffusion")),y=icc,fill=rt_acc))+
  geom_boxplot(position = position_dodge(width=0.75))+
  facet_wrap(~factor(overall_difference, levels=c("overall", "difference"), labels=c("Overall", "Difference")))+
  theme_bw()+
  xlab("")+
  ylab("ICC")+
  theme(legend.title = element_blank())

@@

```{r message=FALSE, warning=FALSE}
tmp = measure_labels %>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)])) %>%
  filter(task == "task") %>%
  select(-var, -task) %>%
  left_join(rel_df[,c("dv", "icc")], by="dv")%>%
  na.omit() %>%
  filter(overall_difference!='condition',
         raw_fit != 'fit',
         rt_acc != 'other')

summary(lm(icc ~ overall_difference*raw_fit,tmp))
```

@@

tmp %>%
  ggplot(aes(days_btw, mean_diff))+
  geom_point()+
  theme_bw()+
  xlab("Days between completion")+
  ylab("Mean standardized difference between two time points")+
  geom_smooth(method="lm")
```

@@

```{r message=FALSE, warning=FALSE}
# boot_df <- read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/input/bootstrap_merged.csv')

boot_df <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_09-27-2017/bootstrap_merged.csv')

@@

```{r warning=FALSE, message=FALSE}
tmp = rel_df %>%
  mutate(var_subs_pct = var_subs/(var_subs+var_ind+var_resid)*100,
         var_ind_pct = var_ind/(var_subs+var_ind+var_resid)*100, 
         var_resid_pct = var_resid/(var_subs+var_ind+var_resid)*100) %>%
  select(dv, task, var_subs_pct, var_ind_pct, var_resid_pct) %>%
  # gather(key, value, -dv, -task) %>%
  mutate(#key = factor(key, levels = c("var_subs_pct", "var_ind_pct", "var_resid_pct")),
         dv = factor(dv, levels = dv[order(task)])) %>%
  separate(dv, c("task_group", "var"), sep="\\.",remove=FALSE,extra="merge") %>%
  mutate(task_group = factor(task_group, levels = task_group[order(task)])) %>%
  arrange(task_group, var_subs_pct) %>%
  mutate(rank = row_number()) %>%
  arrange(task, task_group, rank) %>%
  gather(key, value, -dv, -task_group, -var, -task, -rank) %>%
  ungroup()%>%
  # separate(var, c("var"), sep="\\.",remove=TRUE,extra="drop") %>%
  mutate(task_group = gsub("_", " ", task_group),
         var = gsub("_", " ", var)) %>%
         # var=gsub(".ReflogTr", "", var),
         # var=gsub(".logTr", "", var)) %>%
  mutate(task_group = ifelse(task_group == "psychological refractory period two choices", "psychological refractory period", ifelse(task_group == "angling risk task always sunny", "angling risk task",task_group))) %>%
  mutate(task_group = gsub("survey", "", task_group)) %>%
  filter(task=="task",
         !grepl("EZ|hddm", dv))%>%
  arrange(task_group, rank)
labels = tmp %>%
  distinct(dv, .keep_all=T)

@@

ggsave('Variance_Breakdown_Plot.jpg', plot = p3, device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures", width = 24, height = 20, units = "in")
rm(tmp, labels)
```

@@

```{r}
tmp %>%
  group_by(task, key) %>%
  ggplot(aes(factor(task, levels=c("task","survey"), labels=c("Task", "Survey")), value, fill=key, color=task))+
  geom_boxplot()+
  theme_bw()+
  theme(legend.title = element_blank())+
  scale_fill_manual(breaks = c("var_subs_pct", "var_ind_pct", "var_resid_pct"),
                    labels = c("Variance between individuals",
                               "Variance between sessions",
                               "Error variance"),
                    values=c("grey65", "grey45", "grey25"))+
  scale_color_discrete(guide = FALSE)+
  xlab("")+
  ylab("Percent")

ggsave('Variance_Breakdown_Plot_Summary.jpg', device = "jpeg", path = "/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/figures", width = 10, height = 5, units = "in")
```
